Compositional Verification of Concurrent C Programs with Search Structure Templates

### 2023-05-02 concurrency/CONCUR 2023/paper.tex
unified inFP and Node_ref

\newcommand{\TODO}[1]{\textbf{\textcolor{red}{[ TODO: #1]}}}
%\newcommand{\boxdotright}{\!\mathrel\boxdot\joinrel\rightarrow\!}
\newcommand{\islock}{\boxdotright}
\newcommand{\lockvar}{\islock}
\newcommand{\isaex}{\!\mathrel\odot\joinrel\rightarrow\!}
\newcommand{\xisaex}[1]{\!\mathrel\odot\joinrel\xrightarrow{#1}\!}
%% \newcommand{\ifthenelse}[3]{\text{if }#1\text{ then }#2\text{ else }#3}
\newcommand{\emp}{\mathsf{emp}}

\newcommand\dboxed[1]{\dbox{\ensuremath{#1}}}
\newcommand{\master}[2]{\ensuremath{\mathrm{Master}_{#1}(#2)}}
\newcommand{\snap}[1]{\ensuremath{\mathrm{Snapshot}(#1)}}
\newcommand{\ghost}[2]{\ensuremath{\dboxed{#1}^{#2}}}
\newcommand{\us}{$\mu$s}
\newcommand{\gnamety}{\ensuremath{\mathsf{gname}}}
\newcommand{\treerep}{\ensuremath{\mathsf{Node}}}
\newcommand{\nodeboxrep}{\ensuremath{\mathsf{Node\_ref}}}
\newcommand{\lockinv}{\ensuremath{\mathsf{lock\_inv}}}
\newcommand{\infp}{\ensuremath{\mathsf{InFP}}}

@@

The definitions of $\nodeboxrep$ and $\treerep$ may vary depending on the data structure, but fundamentally, $\nodeboxrep$ holds the resources owned by each client thread, while $\treerep$ holds the abstract representation of the data structure, treated as an atomically accessed shared resource. In particular, holding $\nodeboxrep(\texttt{n})$ should allow us to acquire the lock on node \texttt{n}, which in turn gives us access to the contents of the node. Formally, we write
\[\nodeboxrep(\texttt{n}) \triangleq\ \texttt{n->lock} \lockvar \mathsf{R}(\texttt{n})\]
where the \emph{lock invariant assertion} $\ell \lockvar \mathsf{R}$ expresses that the lock $\ell$ protects resources $\mathsf{R}$. In this case, $\mathsf{R}$ contains the contents of the node (i.e., the \lstinline{node} struct contained inside the \lstinline{node_t} struct of \lstinline{n}).

@@

The local precondition of \lstinline{traverse} includes both the node handle $\nodeboxrep(\texttt{n})$ and its contents $\mathsf{R}(\texttt{n})$. The output of \lstinline{traverse} is a new node \lstinline{n'} in the data structure such that the key \texttt{k} falls within the range of \lstinline{n'}. The local postcondition then includes the handle of the new node $\nodeboxrep(\texttt{n'})$, its contents $\mathsf{R}(\texttt{n'})$, and a Boolean variable $\mathit{res}$ indicating whether \texttt{n'->t} is \texttt{NULL} or not. The variable $\mathit{res}$ is $\mathsf{true}$ if $\texttt{flag}$ returns 1, and $\mathsf{false}$ if 0, mirroring its Boolean state. The key to the correctness of the \texttt{traverse} function is the loop invariant for the top-level loop, which expresses that in each iteration, \lstinline{traverse} holds the lock on a node that has \lstinline{k} in its range:
\begin{align*}\mathsf{traverse\_inv}(\texttt{pn}) \triangleq\ \exists \ \texttt{p}, \texttt{n'}, \mathsf{k'}, \mathsf{v'}, \mathsf{range'}.\ &\texttt{pn} \mapsto (\texttt{p, n'}) \ast \texttt{k} \in \mathsf{range'}\ \ast \  \nodeboxrep(\texttt{n'})  \ \ast \\  &\mathsf{node\_contents}(\texttt{n'}, (\mathsf{k'}, \mathsf{v'}), \mathsf{range'})
\end{align*}

@@

\wm{Should we get rid of inFP and just talk about Node\_ref?}
The give-up template's \texttt{traverse} specification, shown in Figure \ref{fig:traverse_giveup},  is simpler than that of the lock-coupling template. The local precondition does not hold any locks, consisting only of $\infp(\texttt{n})$, a footprint predicate asserting that node \texttt{n} is included in the search structure's abstract state $m$. It is written formally as
$\nodeboxrep(\texttt{n}) \triangleq\ \infp(\texttt{n})$.
%\wm{Should we unify $\infp$ and $\nodeboxrep$?} 
As before, the \texttt{traverse} function will navigate the structure and return a node \texttt{n'} whose range includes \texttt{k}, as well as acquiring \texttt{n'}'s lock and returning its contents $\mathsf{R}(\texttt{n'})$ \wm{or node\_contents?}. Also as before, the Boolean $\mathit{res}$ indicates whether \texttt{n'->t} is \texttt{NULL} or not.

>>>>

\newcommand{\TODO}[1]{\textbf{\textcolor{red}{[ TODO: #1]}}}
%\newcommand{\boxdotright}{\!\mathrel\boxdot\joinrel\rightarrow\!}
%\newcommand{\islock}{\boxdotright}
\newcommand{\lockvar}{\islock}
\newcommand{\isaex}{\!\mathrel\odot\joinrel\rightarrow\!}
\newcommand{\xisaex}[1]{\!\mathrel\odot\joinrel\xrightarrow{#1}\!}
%% \newcommand{\ifthenelse}[3]{\text{if }#1\text{ then }#2\text{ else }#3}
\newcommand{\emp}{\mathsf{emp}}

\newcommand\dboxed[1]{\dbox{\ensuremath{#1}}}
\newcommand{\master}[2]{\ensuremath{\mathrm{Master}_{#1}(#2)}}
\newcommand{\snap}[1]{\ensuremath{\mathrm{Snapshot}(#1)}}
\newcommand{\ghost}[2]{\ensuremath{\dboxed{#1}^{#2}}}
\newcommand{\us}{$\mu$s}
\newcommand{\gnamety}{\ensuremath{\mathsf{gname}}}
\newcommand{\treerep}{\ensuremath{\mathsf{Node}}}
\newcommand{\nodeboxrep}{\ensuremath{\mathsf{Node\_ref}}}
\newcommand{\lockinv}{\ensuremath{\mathsf{lock\_inv}}}
\newcommand{\infp}{\ensuremath{\mathsf{Node\_ref}}}

@@

The definitions of $\nodeboxrep$ and $\treerep$ may vary depending on the data structure, but fundamentally, $\nodeboxrep$ holds the resources owned by each client thread, while $\treerep$ holds the abstract representation of the data structure, treated as an atomically accessed shared resource. In particular, holding $\nodeboxrep(\texttt{n})$ should allow us to acquire the lock on node \texttt{n}, which in turn gives us access to the contents of the node. Formally, we need to know that the triple \[\left\langle \nodeboxrep(n) \ |\ \treerep\ m \right\rangle\ \texttt{acquire}(n\texttt{->lock})\ \left\langle \nodeboxrep(n) \ast \mathsf{R}(n) \ |\ \treerep\ m\right\rangle\] holds, i.e., $\nodeboxrep(n)$ is sufficient to guarantee that node $n$ is in the abstract state of the data structure and its lock protects associated resources $\mathsf{R}(n)$, the \emph{lock invariant}. 
%In this case, $\mathsf{R}$ contains the contents of the node (i.e., the \lstinline{node} struct contained inside the \lstinline{node_t} struct of \lstinline{n}).

@@

The local precondition of \lstinline{traverse} includes both the node handle $\nodeboxrep(\texttt{n})$ and its contents $\mathsf{R}(\texttt{n})$, indicating that a thread should already hold \lstinline{n}'s lock before calling \lstinline{traverse}. The output of \lstinline{traverse} is a new node \lstinline{n'} in the data structure such that the key \texttt{k} falls within the range of \lstinline{n'}. The local postcondition then includes the handle of the new node $\nodeboxrep(\texttt{n'})$, its contents $\mathsf{R}(\texttt{n'})$, and a Boolean variable $\mathit{res}$ indicating whether \texttt{n'->t} is \texttt{NULL} or not. The variable $\mathit{res}$ is $\mathsf{true}$ if $\texttt{flag}$ returns 1, and $\mathsf{false}$ if 0, mirroring its Boolean state. The key to the correctness of the \texttt{traverse} function is the loop invariant for the top-level loop, which expresses that in each iteration, \lstinline{traverse} holds the lock on a node that has \lstinline{k} in its range:
\begin{align*}\mathsf{traverse\_inv}(\texttt{pn}) \triangleq\ \exists \ \texttt{p}, \texttt{n'}, \mathsf{k'}, \mathsf{v'}, \mathsf{range'}.\ &\texttt{pn} \mapsto (\texttt{p, n'}) \ast \texttt{k} \in \mathsf{range'}\ \ast \  \nodeboxrep(\texttt{n'})  \ \ast \\  &\mathsf{node\_contents}(\texttt{n'}, (\mathsf{k'}, \mathsf{v'}), \mathsf{range'})
\end{align*}

@@

\wm{Should we get rid of inFP and just talk about Node\_ref?}
The give-up template's \texttt{traverse} specification, shown in Figure \ref{fig:traverse_giveup}, is slightly simpler than that of the lock-coupling template: a thread does not need to hold any locks before calling it, so the invariant $\mathsf{R}$ does not appear in the precondition.
As before, the \texttt{traverse} function will navigate the structure and return a node \texttt{n'} whose range includes \texttt{k}, as well as acquiring \texttt{n'}'s lock and returning its contents $\mathsf{R}(\texttt{n'})$ \wm{or node\_contents?}. Also as before, the Boolean $\mathit{res}$ indicates whether \texttt{n'->t} is \texttt{NULL} or not.
